Hamiltonian mechanics is conservation of information entropy

### 2018-10-26 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
First draft probably done

\begin{abstract}
	In this work we show that the canonical transformations of classical Hamiltonian mechanics are exactly the transformations that preserve information.
\end{abstract}

@@

@@

The link to information theory is the following: suppose we have a distribution, known to both of us, and you select one element at random. How many yes or no questions do you have to answer before I know what case you selected? For example, if you have a pouch with 230 balls of which 115 are red and 115 are green, one yes or no question would be sufficient. In general, if $I$ is the number of questions then $C = 2 ^ I$ are the number of cases you can distinguish with those questions. But suppose $\rho(\xi_0)=1/5$, then we know that we'd select $\xi_0$ one in 5 cases. That is, the normalized distribution is be the inverse of the number of cases: $\rho(\xi_0) = \frac{1}{C}$.  Therefore the information needed to identify a particular element is $I(\xi^a)=\log \frac{1}{\rho(\xi^a)}$. If we take the average and change the base of the logarithm, we have:
\begin{equation}
I(\rho(\xi^a)) = \int \rho(\xi^a)) \ln \left(\frac{1}{\rho(\xi^a))}\right) d\xi^1 d\xi^2 ... d\xi^n =-\int \rho(\xi^a)) \ln (\rho(\xi^a)) d\xi^1 d\xi^2 ... d\xi^n
\end{equation}
which is the information entropy of the distribution.\footnote{The nature of information entropy is often misunderstood, especially in physics. It does not represent the information the distribution has: it represent the \emph{additional} information that would be needed to identify an element. Also, it does not represent knowledge: as we are studying actual physical distributions, any correct distribution has full knowledge of the system.}

@@

Similarly, we can see that information entropy is not invariant
\begin{align*}
I(\rho(\xi^b)) &=-\int \rho(\xi^b) \ln (\rho(\xi^b)) d\xi^1 ... d\xi^n \\
&=-\int \rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| \ln \left(\rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\|\right) d\xi^1 ... d\xi^n \\
&=-\int \rho(\hat{\xi}^a) \ln \left(\rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\|\right) d\hat{\xi}^1 ... d\hat{\xi}^n \\
&=-\int \rho(\hat{\xi}^a) \ln (\rho(\hat{\xi}^a)) d\hat{\xi}^1 ... d\hat{\xi}^n -\int \rho(\hat{\xi}^a) \ln \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| d\hat{\xi}^1 ... d\hat{\xi}^n
\end{align*}
\begin{equation}\label{entropy_transformation}
I(\rho(\xi^b)) =I(\rho(\hat{\xi}^a)) -\int \rho(\hat{\xi}^a) \ln \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| d\hat{\xi}^1 ... d\hat{\xi}^n
\end{equation}
The reason is that we may be changing scale and therefore level of precision.

The situation is actually worse. We'd like to think of a density as a number associated to a point of a space but that's technically incorrect. Given a manifold $\mathcal{M}$ upon which we define the distribution, the latter is \emph{not} $\rho : \mathcal{M} \to \mathbb{R}$. That would be invariant change of variables, as a point is a point no matter how is it identified. Instead we additionally need a set of functions $\xi^b : \mathcal{M} \to \mathbb{R}$ and only then we can define the density as $\rho : \mathbb{R}^n \to \mathbb{R}$. The density value is defined on the values for a set of variables, not at a point.\footnote{As the density is really a limit, we need to know how that limit is taken and the coordinates decide that.} That is, the density is undefined without a (differentiable) set of coordinates.

@@

In probability theory one starts with three objects: a set of outcomes $\mathcal{M}$, a set of events $\sigma_\mathcal{M}$ where each event is a collection of outcomes, and a probability measure $\mu : \sigma_\mathcal{M} \rightarrow [0,1]$ that assigns a number between zero and one to each event. That is, we don't assign a probability density to the points of $\mathcal{M}$ but we assign a finite probability to sets of points. We can then define a random variable $\xi : \mathcal{M} \rightarrow \mathbb{R}$ as a real valued function of the outcomes. For each random variable we can define its cumulative distribution function $F_\xi(\xi_0)=\mu(\xi<\xi_0)$ and the probability density function is its Radon-–Nikodym derivative $\rho(\xi) = \frac{dF_\xi}{d\xi}$. That is: the integral is the primal object while the density is derived.

@@

In differential geometry, one starts with a manifold: a set of points $\mathcal{M}$ which can be given local coordinates $\xi^i : \mathcal{M} \to \mathbb{R}$. One defines a tangent space $\mathsf{T}\mathcal{M}$ where vectors live and cotangent space $\mathsf{T}^*\mathcal{M}$ were linear functions of vectors live. Then we define $n$-forms as multi-linear functions of $n$ vectors that return the value associated with the infinitesimal parallelepiped they form. That is $\nu : (\mathsf{T}^*\mathcal{M})^n \rightarrow \mathbb{R}$. As vectors are coordinate invariant, $n$-forms are coordinate invariant as well so we can write $m_V=\int_V \nu(dV)$ with no reference to coordinates. If $e_i$ are the basis vector associated with $\xi^i$ and $e^j$ are linear functions such that $e^j(e_i)=\delta_i^j$, then we can express each $n$-form as $\rho(\xi^i)e_1\wedge e_2 \wedge ... \wedge e_n$. The density $\rho(\xi^i)$ is the component of the $n$-form expressed in the $\xi^i$ coordinates in the a way analogous to components of a vector.\footnote{Technically, the form is linear so it changes with the Jacobian determinant and not its absolute value. But negative determinants means changing orientation of the region of integration, so fixing that would bring another sign change and $\rho(q^i)$ would not change. This is equivalent to saying that some coordinate systems are right handed and some are left handed.}

This approach gives us a way to understand which objects are coordinate independent and which aren't. But it has two issues: it does not allow to easily talk about coordinate dependent operations, such as marginal distributions and information entropy, and it may use definitions that do not map well to physical concepts.\footnote{Typically, a vector is defined as a map between a scalar function of the manifold to another scalar function to the manifold. Or as an equivalence class of trajectories. It is a stretch to think of fluid velocities or electric fields in those terms.}

So the situation is that we have a single clear physical objects we want to study, but different and somewhat disconnected mathematical frameworks to study it. This should not be surprising as the frameworks are defined by mathematicians to solve their own needs, which is totally understandable. But it is an issue as each mathematical framework only partially captures our physical understanding and we risk missing the full picture.

\section{Coordinate independent densities}

@@

If the evolution is deterministic and reversible, then all the parts that are in one state will be mapped to another: the density should be exactly mapped to a different state and therefore should not change value. If the evolution is deterministic and reversible, then if I give you enough information to identify an element at one time I also gave you enough information to identify the element at a future time: the information entropy will remain the same over time. And here lies the problem: how can we make sure that these quantities are preserved over time if they are not even observer independent? How can we know if the density stays the same if we can't even compare the value of the densities at two different elements of the distribution? How can we tell whether the information entropy is conserved if we can't event define a unique value at a given time?

\textbf{If we want to define deterministc and reversible evolution over distributions, they must be invariant distributions. That is, the densities and information entropy must be independent on the choice of coordinates.}

@@

Suppose $\mathcal{S}$ is a two dimensional manifold. Suppose we have two variables $(q,k)$ such that $k$ uses the inverse units of $q$. For example, if $q$ is meters then $k$ is inverse meters and an infinitesimal area $dq dk$ will be a pure number. We say $q$ is a coordinate as it is a variable that defines a unit, while $k$ is the corresponding conjugate variable as it uses the inverse unit. Now suppose you change coordinate. For example, $\hat{q}$ will be kilometers and $\hat{k}$ inverse kilometers. We'll have $d\hat{q} d\hat{k} = dq dk$. Therefore we'll also have $\rho(q,k) = \rho(\hat{q}, \hat{k})$: the density is invariant and so is the information entropy.

@@

@@

Let's set $\hbar$ as the unit we use to describe the amount of possible states. We have:
\begin{equation}
m_V = M \mu(V) =\int_V M \rho(q, k) \hbar dq dk = \int_V M \rho(q, p) dq dp
\end{equation}
where we set $p=\hbar k$. We should recognize $p$ as conjugate momentum and $k$ the classical analogue of the wave number. Suppose we change variables. In general the Jacobian will be given by:
\begin{equation}
\label{Poisson}
\begin{aligned}
|J| &= \left| \begin{matrix}
\dfrac{\partial \hat{q}}{\partial q} & \dfrac{\partial \hat{q}}{\partial p} \\[2.2ex]
\dfrac{\partial \hat{p}}{\partial q} & \dfrac{\partial \hat{p}}{\partial p} \end{matrix} \right| = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} - \frac{\partial \hat{p}}{\partial x} \frac{\partial \hat{q}}{\partial p} &= \{\hat{q}, \hat{p}\}
\end{aligned}
\end{equation}
which we recognize it to be the Poisson bracket. If the change of variable is simply a change of coordinates, a change of units, we have:
\begin{equation}
\label{coordinate_change}
\begin{aligned}
\hat{q} &= \hat{q}(q) \\
\{\hat{q}, \hat{p}\} &= 1 = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} \\
\dfrac{\partial \hat{p}}{\partial p} &= \frac{\partial \hat{q}}{\partial q} ^{-1} = \frac{\partial q}{\partial \hat{q}} \\
\hat{p} &= \frac{\partial q}{\partial \hat{q}} p
\end{aligned}
\end{equation}
Conjugate momentum transforms as a covariant component. Under coordinate transformations we have:
\begin{equation}
\label{density_invariance}
\begin{aligned}
\rho(q,p) &= \rho(\hat{q}, \hat{p}) \\
I(\rho(q,p)) &= I(\rho(\hat{q},\hat{p}))
\end{aligned}
\end{equation}
Let us call $(q,p)$ canonical variables if the density is equal to the invariant density. Then two other variables such that $\{\hat{q}, \hat{p}\}=1$ are also canonical since the Jacobian determinant will be unitary.

@@

We'd like to stress how all elements of Hamiltonian mechanics had a straight forward physical meaning in terms of invariant densities and change of coordinates and variables.

@@

@@

While the power of this shift can really be appreciated in the context of our broder work, it should be clear how with relative few starting points, elements of relativity (e.g. coordinate invariance), statistical mechanics (e.g. physical distributions), quantum mechanics (e.g. wave number, conjugate quantities) and classical Hamiltonian mechanics (e.g. phase space, Poisson brackets) already emerge and form a coherent picture. In particular note how energy (in the form of the Hamiltonian) and entropy (in the form of information entropy) both come out from the single unassuming idea of invariant distributions. And note how conservation of entropy is related to deterministic and reversible evolution. Also note how deterministic and reversible evolution means the that the system is only influenced by itself, which means it is not influenced by anything else, which means it is isolated, which indeed means energy is conserved. This surely this cannot be all coincidence.

@@

In the narrower context, it allows to understand what physical object is represented by each mathematical one. This can be summarized in table \ref{dictionary}.
\begin{table}[h]
	\centering
	%	\begin{tabular}{p{0.2\textwidth} p{0.1\textwidth} p{0.1\textwidth} p{0.5\textwidth}}
	\begin{tabular}{c p{0.3\textwidth} p{0.5\textwidth} }
		& Name & Meaning\\ 
		\hline 
		& Particle (classical) & an infinitesimal part of the system (i.e. the limit of recursive division) \\ 
		$\mathcal{S} =\mathsf{T}^*\mathcal{Q}$ & Phase space \newline (Tangent bundle) & the set of all possible states for particles upon which invariant distributions can be defined \\
		$\xi^a$ & State variables \newline (Unified coordinates) & the set of variables needed to identify the state of a particle \\ 
		$\rho(\xi^a)$ & Density & the amount of material for a given particle state (i.e. in the limit of the recursive division)\\ 
		$I(\rho(\xi^a))$ & Information entropy & the number of bits to identify an element of the distribution to a unitary level of precision\\ 
		$q^i$ & Coordinate & a state variable that also defines a unit \\
		$\mathcal{Q}$ & Coordinate space \newline (Configuration space) & the space charted by all coordinates \\
		$k_i$ & Conjugate coordinate & a state variable that uses the inverse unit of the corresponding coordinate \\
		$\hbar$ & & the unit for measuring the range of possible states within a single degree of freedom \\
		$p_i=\hbar k_i$ & Conjugate momentum & a state variable that together with the corresponding coordinate defines invariant ranges of possible states \\
		$(q^i, p_i)$ & Canonical variables \newline (Canonical coordinates) & a set of variable for which the density is invariant and expressed over units of $\hbar$\\ 
		& Canonical transformation & a change in state variables that does not change the density\\ 
	\end{tabular}
	\caption{Dictionary between mathematical and physical objects.}
	\label{dictionary}
\end{table}
It also allows us to see that deterministic and reversible evolution must conserve entropy and that isolated systems must conserve energy. In the broader context, it allows to make deeper connections between different areas of math and physics, and pushes us to rethink our starting points. In particular, it tells us that classical particles should not be considered point-like objects (upon which entropy and densities cannot be defined), but the limit of recursive subdivisions (i.e. infinitesimal cells of phase space).

@@

>>>>

\begin{abstract}
	In this work we show the equivalence between Hamiltonian mechanics and conservation of information entropy. We will show that distribution with coordinate independent values for information entropy require that the manifold on which the distribution is defined is charted by conjugate pairs (i.e. it is a symplectic manifold). We will also show that further requiring that the information entropy is conserved during the evolution yields Hamilton's equations. This will provide new insights in the nature of classical mechanics and physics in general, as it will connect to elements of relativity, statistical mechanics, thermodynamics, quantum mechanics, probability theory and symplectic geometry. This work is part of a larger project, called Assumptions of physics, which aims to rederive the known equations from a handful of physically meaningful starting points.
\end{abstract}

@@

Over the past decade there has been a renewed interest in the underpinning of classical mechanics~\cite{North,Curiel,Barrett1,Barrett2}. Different attempts and arguments have been put forward to understand what are the most important feature of the different mathematical formulations, whether one is more fundamental of the other in some sense or whether they are equivalent.

All these attempts, though, suffer from one problem. They concentrate on the mathematical structure. The issue is that, in general, a mathematical object can represent more than one physical entity. For example, the same mathematical framework for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems. On the other end, the same physical entity can be described by more than one mathematical object. For example, the same cannonball can be described by position and velocity in a Lagrangian setting, position and momentum in a Hamiltonian setting or in a thermodynamics setting by temperature, volume and so on. The connection between math and physics, then, is many-to-many and therefore we can't study only one side. The reason is that mathematics captures only the syntax of a particular description without its semantics, which at that point cannot be recovered. Mathematical structures are just shadows of the physical model they represent.

It should be clear that we do not fault the authors. Even just a century ago, there was a push to find fundamental principles or laws that could serve as a starting point for the different theories. Newtonian mechanics, thermodynamics and special relativity serve as good examples. Nowadays, unfortunately, much of the current fundamental theories, including Lagrangian, Hamiltonian, Quantum mechanics and Quantum Field Theory take as their starting point a particular mathematical structure. That is, we do not know what the value of a Lagrangian is or what the symplectic form represents: we just take them as a given. Moreover, a lot of work in theoretical physics nowadays starts by exploring novel abstract mathematical structures and then trying to see whether some physical meaning or prediction can be extracted. We are not going to debate here how and why this shift happened. We just merely note that it leaves the semantics of our theories ill defined.\footnote{Which is a nice way to say that we don't really know what are we talking about.}

It is only natural that, given the circumstances, one uses the only thing that seems to be well defined, the math, and tries to get understanding from there. But here lies the fundamental problem: that math is not created by physicists to do physics, it is created by mathematicians to do math. The definitions of topology, differentiable manifold, symplectic manifold, Hilbert space and so on are chosen either because they are convenient to prove theorems or because they are convenient to make calculations.\footnote{In fact, it is common in math to have two such definitions and then prove they are equivalent.} We should not expect them to be a good match for definitions that would be physically meaningful. In other words, there is a very good chance that just looking at just the math may lead us, like horses with blinders, to a very clear path which is not the physically meaningful one.\footnote{To be clear, we do not fault the mathematicians either as they are not responsible for whether their structures are used appropriately or not.}

Fully conscious of these issues, we have embarked in an effort, called Assumptions of Physics~\cite{carc1}, that aims to find a handful of physical principles from which the known laws of physics can be formally derived~\cite{carc2}. The approach has several advantages. First, by starting from physical ideas we put physics back at the center of the discussion, and not relegate it as a mere after the fact interpretation: the semantics is always well defined. Second, it puts physics on sturdier mathematical grounds as we have a precise physical meaning to all the formal objects and we leave out those that do not. Third, it fosters connections between different fields of science, math, philosophy and engineering: it is our belief that nature is one and does not care about such divisions and therefore truly fundamental concepts must span freely across different areas. Fourth, it gives a framework to pose deep questions of the type the philosophy of science community is interested in. And lastly, by being forced to formalize and make explicit our assumptions, we can better understand how and why they fail possibly giving new insights that may lead to new physics. While the project is ambitious, we have already made progress beyond what we thought it could be reasonably done. So there's that.

In this paper we focus on just one particular insight provided by our work: \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e. a symplectic structure where variables are organized in conjugate pairs). In section 2 we show that, in general, information entropy and distributions densities are coordinate dependent. In section 3 we show that the structure of phase space is required to make such objects invariant under coordinate transformation (i.e. observer independent) and therefore give them an objective physical character. Conservation of entropy over time will force that invariance during time evolution, giving rise to Hamilton's equations. In chapter 4 we will briefly discuss the result and show how it necessarily leads to a classical version of the uncertainty principle.

The overall idea is that \textbf{Hamiltonian mechanics is best understood as the deterministic and reversible\footnote{By reversible we mean the ability, given a state, to reconstruct its past. This is the dual of determinism, that is the ability, given a state, to predict its future.} evolution of infinitesimally reducible objects}. As the information needed to identify an element of the distribution does not change in time, knowing the initial state is equivalent to knowing the future or past state. Classical particles, then, should be understood as the limit of subsequent subdivisions, infinitesimally small objects with extent in phase space. This single physical model is captured piecemeal by elements of statistics, probability theory and symplectic geometry which are given a coherent and well defined meaning. In our view, this understanding also meshes better with other ideas from general relativity\footnote{A point particle in general relativity would not follow a geodesic and would, due to the infinite mass density, puncture space-time.}, thermodynamics and quantum mechanics.

@@

The link to information theory is the following: suppose we have a distribution, known to both of us, and you select one element at random. How many yes or no questions do you have to answer before I know what case you selected? For example, suppose there are 4 balls of which 2 are red and 2 are green, one yes or no question would be sufficient. In general, if $I$ is the number of questions then $C = 2 ^ I$ are the number of cases you can distinguish with those questions. On the other hand, suppose $\rho(\xi_0)=1/5$, then we know that we'd select $\xi_0$ one in 5 cases. That is, the normalized distribution is the inverse of the number of cases: $\rho(\xi_0) = \frac{1}{C}$.  Therefore the information needed to identify a particular element is $I(\xi^a)=\log \frac{1}{\rho(\xi^a)}$. If we take the average and change the base of the logarithm, we have:
\begin{equation}
I(\rho(\xi^a)) = \int \rho(\xi^a)) \ln \left(\frac{1}{\rho(\xi^a))}\right) d\xi^1 d\xi^2 ... d\xi^n =-\int \rho(\xi^a)) \ln (\rho(\xi^a)) d\xi^1 d\xi^2 ... d\xi^n
\end{equation}
which is the information entropy of the distribution.\footnote{The nature of information entropy is often misunderstood, especially in physics. It does not represent the information the distribution has: it represent the \emph{additional} information that would be needed to identify an element. Also, it does not represent knowledge: as we are studying actual physical distributions, any correct distribution has full knowledge of the system.}

@@

Similarly, we can see that information entropy is not invariant
\begin{align*}
I(\rho(\xi^b)) &=-\int \rho(\xi^b) \ln (\rho(\xi^b)) d\xi^1 ... d\xi^n \\
&=-\int \rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| \ln \left(\rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\|\right) d\xi^1 ... d\xi^n \\
&=-\int \rho(\hat{\xi}^a) \ln \left(\rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\|\right) d\hat{\xi}^1 ... d\hat{\xi}^n \\
&=-\int \rho(\hat{\xi}^a) \ln (\rho(\hat{\xi}^a)) d\hat{\xi}^1 ... d\hat{\xi}^n -\int \rho(\hat{\xi}^a) \ln \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| d\hat{\xi}^1 ... d\hat{\xi}^n
\end{align*}
\begin{equation}\label{entropy_transformation}
I(\rho(\xi^b)) =I(\rho(\hat{\xi}^a)) -\int \rho(\hat{\xi}^a) \ln \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| d\hat{\xi}^1 ... d\hat{\xi}^n
\end{equation}
The reason is that we may be changing scale and therefore level of precision. Given that both change is determined by the Jacobian, we conclude that \textbf{during a change of variable the information entropy of all possible distributions does not change if and only if the density remains the same at every point under the same change of variable}.

To sum up, while we'd like to think of a density as a number associated to a point of a space, that's technically incorrect. Given a manifold $\mathcal{M}$ upon which we define the distribution, the latter is \emph{not} $\rho : \mathcal{M} \to \mathbb{R}$. That would be a scalar function, invariant under change of variables, as a point is a point no matter how is it identified. Instead we additionally need a set of functions $\xi^a : \mathcal{M} \to \mathbb{R}$ and only then we can define the density as $\rho : \mathbb{R}^n \to \mathbb{R}$. The density value is defined on the values for a set of variables, not at a point.\footnote{As the density is really a limit, we need to know how that limit is taken and the coordinates decide that.} That is, the density is undefined without a (differentiable) set of coordinates.

@@

In probability theory one starts with three objects: a set of outcomes $\mathcal{M}$, a set of events $\sigma_\mathcal{M}$ where each event is a collection of outcomes, and a probability measure $\mu : \sigma_\mathcal{M} \rightarrow [0,1]$ that assigns a number between zero and one to each event. That is, we don't assign a probability density to the points of $\mathcal{M}$ but we assign a finite probability to finite regions. We can then define a random variable $\xi : \mathcal{M} \rightarrow \mathbb{R}$ as a real valued function of the outcomes. For each random variable we can define its cumulative distribution function $F_\xi(\xi_0)=\mu(\xi<\xi_0)$ and the probability density function is its Radon-–Nikodym derivative $\rho(\xi) = \frac{dF_\xi}{d\xi}$. That is: the integral is the primal object while the density is derived.

@@

In differential geometry, one starts with a manifold: a set of points $\mathcal{M}$ which can be given local coordinates $\xi^i : \mathcal{M} \to \mathbb{R}$. One defines a tangent space $\mathsf{T}\mathcal{M}$ where vectors live and cotangent space $\mathsf{T}^*\mathcal{M}$ were linear functions of vectors live. Then we define $n$-forms as multi-linear functions of $n$ vectors that return the value associated with the infinitesimal parallelepiped they form. That is $\nu : (\mathsf{T}^*\mathcal{M})^n \rightarrow \mathbb{R}$. As vectors are coordinate invariant, $n$-forms are coordinate invariant as well so we can write $\mu(V)=\int_V \nu(dV)$ with no reference to coordinates. If $e_a$ are the basis vector associated with $\xi^a$ and $e^b$ are linear functions such that $e^b(e_a)=\delta_a^b$, then we can express each $n$-form as $\rho(\xi^a)e_1\wedge e_2 \wedge ... \wedge e_n$. The density $\rho(\xi^a)$ is the component of the $n$-form expressed in the $\xi^a$ coordinates in the a way analogous to components of a vector.\footnote{Technically, the form is linear so it changes with the Jacobian determinant and not its absolute value. But negative determinants means changing orientation of the region of integration, so fixing that would bring another sign change and $\rho(\xi^a)$ would not change. This is equivalent to saying that some coordinate systems are right handed and some are left handed.}

This approach gives us a way to understand which objects are coordinate independent and which aren't. But it has two issues: it does not allow to easily talk about coordinate dependent operations, such as marginal distributions and information entropy, and it may use definitions that do not map well to physical concepts.\footnote{Typically, a vector is defined as a map between a scalar function of the manifold to another scalar function to the manifold. Or as an equivalence class of trajectories. It is a stretch to think of fluid velocities or electric fields in those terms. For that same reason, we do not use the differential geometry notation $\frac{\partial}{\partial \xi^a}$ and $d\xi^a$ for vector and covector basis because they do not convey the actual physical meaning of these objects as they are used in physics.}

So the situation is that we have a single clear physical objects we want to study, but different and somewhat disconnected mathematical frameworks to study it. This should not be surprising as the frameworks are defined by mathematicians, as we said before, to solve their own needs. But it is an issue when these are used in physics without careful consideration as each mathematical framework only partially captures our physical understanding. That is why we argue that simply examining the mathematical structure used in a physical theory is not enough to gain a complete picture.

\section{Coordinate independent distributions}

@@

If the evolution is deterministic and reversible, then all the parts that are in one state will be mapped to another: the density should be exactly mapped to a different state and therefore should not change value. If the evolution is deterministic and reversible, then if I give you enough information to identify the state of an element at one time I also gave you enough information to identify the state of the same element at a future time: the information entropy will remain the same over time. And here lies the problem: how can we make sure that these quantities are preserved over time if they are not even observer independent? How can we know if the density stays the same if we can't even compare the value of the densities at two different elements of the distribution? How can we tell whether the information entropy is conserved if we can't event define a unique value at a given time?

This leads us to the following insight: \textbf{if we want to define deterministc and reversible evolution over distributions, they must be invariant distributions. That is, the densities and information entropy must be independent on the choice of coordinates.} As we noted before, requiring one to be coordinate invariant will automatically require the other to be invariant as well.

@@

Suppose $\mathcal{S}$ is a two dimensional manifold. Suppose we have two variables $(q,k)$ such that $k$ uses the inverse units of $q$. For example, if $q$ is meters then $k$ is inverse meters which means an infinitesimal area $dq dk$ will be a pure number. We say $q$ is a coordinate as it is a variable that defines a unit, while $k$ is the corresponding conjugate variable as it uses the inverse unit. Now suppose you change coordinate. For example, $\hat{q}$ will be kilometers and $\hat{k}$ inverse kilometers. We'll have $d\hat{q} d\hat{k} = dq dk$. Therefore we'll also have $\rho(q,k) = \rho(\hat{q}, \hat{k})$: the density is invariant and so is the information entropy.

@@

Physically, the coordinate plays a double role. It both defines a unit and it helps identify particle states. It is this double role that gives it an important character because a change in $q$ must induce a change in $k$ as their units and the unit of $\rho$ are not independent. Symplectic geometry strips away the units and does not elevate densities as primary objects. It is no wonder the link gets lost.

@@

Let's set $\hbar$ as the unit we use to describe the amount of possible states (or the inverse of a unitary density). We have:
\begin{equation}
m_V = M \mu(V) =\int_V M \rho(q, k) \hbar dq dk = \int_V M \rho(q, p) dq dp
\end{equation}
where we set $p=\hbar k$. We should recognize $p$ as conjugate momentum and $k$ the classical analogue of the wave number. Suppose we change variables. In general the Jacobian will be given by:
\begin{equation}
\label{Poisson}
\begin{aligned}
|J| &= \left| \begin{matrix}
\dfrac{\partial \hat{q}}{\partial q} & \dfrac{\partial \hat{q}}{\partial p} \\[2.2ex]
\dfrac{\partial \hat{p}}{\partial q} & \dfrac{\partial \hat{p}}{\partial p} \end{matrix} \right| = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} - \frac{\partial \hat{p}}{\partial x} \frac{\partial \hat{q}}{\partial p} &= \{\hat{q}, \hat{p}\}
\end{aligned}
\end{equation}
which we recognize it to be the Poisson bracket. If the change of variable is simply a change of coordinates, a change of units, we have:
\begin{equation}
\label{coordinate_change}
\begin{aligned}
\hat{q} &= \hat{q}(q) \\
\{\hat{q}, \hat{p}\} &= 1 = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} \\
\dfrac{\partial \hat{p}}{\partial p} &= \frac{\partial \hat{q}}{\partial q} ^{-1} = \frac{\partial q}{\partial \hat{q}} \\
\hat{p} &= \frac{\partial q}{\partial \hat{q}} p
\end{aligned}
\end{equation}
Conjugate momentum transforms as a covariant component. Under coordinate transformations we have:
\begin{equation}
\label{density_invariance}
\begin{aligned}
\rho(q,p) &= \rho(\hat{q}, \hat{p}) \\
I(\rho(q,p)) &= I(\rho(\hat{q},\hat{p}))
\end{aligned}
\end{equation}
Let us call $(q,p)$ canonical variables if the density is equal to the invariant density. Then two other variables such that $\{\hat{q}, \hat{p}\}=1$ are also canonical since the Jacobian determinant will be unitary.

@@

We'd like to stress how, once the simple premise of invariant distributions is accepted, all elements of Hamiltonian mechanics are necessary and have a straight forward physical meaning. Not only that: because of the equivalent we know that everything in Hamiltonian mechanics can be understood with just that simple premise. And the premise is not about cotangent bundles of Lie group of quasi-q spaces on an infinite-dimensonal lattice. It's just the simple realization that if we require distributions to have a physically objective reality they must be coordinate independent.

@@

The symplectic structure was recovered by requiring invariant distributions which in turns requires invariant integration over independent degrees of freedom. Note how the requirement was given outside of the symplectic structure itself and therefore it cannot be recovered simply from that.

@@

Also note that this view tells us exactly what the mathematical framework of symplectic geometry is supposed to represent. The state space of particles is a manifold because we use real numbers to identify the states. Among all possible choice of state variables, we conveniently choose the ones in which the distribution can be expressed as a density. That's the differentiable structure on the manifold. The tangent and cotangent space are the tools to keep track of the process of infinitesimal subdivision and they are linear spaces because the space of all possible distributions is itself a linear space. Moreover the differentiable manifold has a symplectic structure because the density is invariant under coordinate transformation. Therefore among all possible choice of differentiable state variable we choose the canonical ones, the ones in which the density is expressed in appropriate units. Every requirement is physically motivated. No extra requirements are left.

While the power of this shift can really be appreciated in the context of our broder work, it should be clear how with relative few starting points, elements of relativity (e.g. coordinate invariance), statistical mechanics (e.g. physical distributions), quantum mechanics (e.g. wave number, conjugate quantities) and classical Hamiltonian mechanics (e.g. phase space, Poisson brackets) already emerge and form a coherent picture. In particular note how energy (in the form of the Hamiltonian) and entropy (in the form of information entropy) both come out from the single unassuming idea of invariant distributions. And note how conservation of entropy is related to deterministic and reversible evolution. Also note how deterministic and reversible evolution means the that the system is only influenced by itself, which means it is not influenced by anything else, which means it is isolated, which it's linked to energy conservation.\footnote{To be precise, conservation of energy is not an invariant property. The fact that a Hamiltonian exists, though, is an invariant property.} This surely this cannot be all coincidence.

@@

In the narrower context, it allows to understand what physical object is represented by each mathematical one. This can be summarized in table \ref{dictionary}.
\begin{table}[h]
	\centering
	%	\begin{tabular}{p{0.2\textwidth} p{0.1\textwidth} p{0.1\textwidth} p{0.5\textwidth}}
	\begin{tabular}{c p{0.3\textwidth} p{0.5\textwidth} }
		& Name & Meaning\\ 
		\hline 
		& Classical particle & an infinitesimal part of the system (i.e. the limit of recursive division) \\ 
		$\mathcal{S} =\mathsf{T}^*\mathcal{Q}$ & Phase space \newline (Tangent bundle) & the set of all possible states for particles upon which invariant distributions can be defined \\
		$\xi^a$ & State variables \newline (Unified coordinates) & the set of variables needed to identify the state of a particle \\ 
		$\rho(\xi^a)$ & Density & the amount of material for a given particle state (i.e. in the limit of the recursive division)\\ 
		$I(\rho(\xi^a))$ & Information entropy & the number of bits to identify an element of the distribution to a unitary level of precision\\ 
		$q^i$ & Coordinate & a state variable that also defines a unit \\
		$\mathcal{Q}$ & Coordinate space \newline (Configuration space) & the space charted by all coordinates \\
		$k_i$ & Conjugate coordinate & a state variable that uses the inverse unit of the corresponding coordinate \\
		$\hbar$ & & the unit for measuring the range of possible states within a single degree of freedom \\
		$p_i=\hbar k_i$ & Conjugate momentum & a state variable that together with the corresponding coordinate defines invariant ranges of possible states \\
		$(q^i, p_i)$ & Canonical variables \newline (Canonical coordinates) & a set of variable for which the density is invariant and expressed over units of $\hbar$\\ 
		& Canonical transformation & a change in state variables that does not change the density\\ 
	\end{tabular}
	\caption{Dictionary between mathematical and physical objects.}
	\label{dictionary}
\end{table}
It also allows us to see that deterministic and reversible evolution must conserve entropy and that isolated systems must conserve energy. In the broader context, it allows to make deeper connections between different areas of math and physics, and pushes us to rethink our starting points. In particular, it tells us that classical particles should not be considered point-like objects (upon which entropy and densities cannot be defined), but the limit of recursive subdivisions (i.e. infinitesimal cells of phase space).

@@

	\bibitem{North} Jill North; The "Structure" of Physics: A Case Study, The Journal of Philosophy 106, no. 2 (2009), Pages 57-88, http://www.jstor.org/stable/20620153.

	\bibitem{Curiel}Erik Curiel; Classical Mechanics Is Lagrangian; It Is Not Hamiltonian, The British Journal for the Philosophy of Science, Volume 65, Issue 2, 1 June 2014, Pages 269–321, https://doi.org/10.1093/bjps/axs034

	\bibitem{Barrett1}Thomas William Barrett; On the Structure of Classical Mechanics, The British Journal for the Philosophy of Science, Volume 66, Issue 4, 1 December 2015, Pages 801–828, https://doi.org/10.1093/bjps/axu005

	\bibitem{Barrett2}Thomas William Barrett; Equivalent and Inequivalent Formulations of Classical Mechanics, The British Journal for the Philosophy of Science, , axy017, https://doi.org/10.1093/bjps/axy017

	\bibitem{carc1} G. Carcassi, C. A. Aidala: \emph{Assumptions of Physics} (in preparation). http://assumptionsofphysics.org/book/, Ann Arbor, MI, USA, 2018.

	\bibitem{carc2} G. Carcassi, C. A. Aidala, D. J. Baker and L. Bieri; From physical assumptions to classical and quantum Hamiltonian and Lagrangian particle mechanics, Journal of Physics Communications, 2, 4, 045026, 2018.
